The Success of College Basketball Teams can be predicted using several advanced team statistics

### 2022-04-22 Paper-On-NCAAB-Stats/outputs/paper/Paper-On-NCAAB-Stats.Rmd
data visualizations

library(tidyverse)
library(dplyr)
library(knitr)
```

@@

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

>>>>

library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggpubr)
library(modelsummary)
```

@@

```{r dataextract, echo = FALSE, message=FALSE}
#Load in dataset
ncaab_stats <- read_csv(here::here("inputs/data/ncaab_stats.csv"))

#Show a brief extract of the dataset by selecting 10 observations in a table
ncaab_stats %>% 
  slice(20,112,630,702,785,791,1107,1221,1493,1558) %>%
  kable(
    caption= "10 Observations from dataset of NCAA Basketball team statistics",
    col.names = c("Year", "School", "Over .500", "SRS", "Pace", "O-rating", "3pa Rate", "TS %", "TRB %", "AST %", "STL %", "BLK %", "TOV %", "ORB %"),
    booktabs = TRUE,
    linesep = "",
    align = c('c')
  ) %>% 
  kable_styling(latex_options="scale_down")

```

```{r srshist, echo = FALSE, message=FALSE, fig.cap= "Distribution of Simple Rating System for NCAA Basketball teams"}
ncaab_stats %>%
  ggplot(aes(x = srs)) + 
  geom_histogram(bins = 50, fill = "dodgerblue1", colour = "black") +
  theme(legend.position = "none")+
  labs(x = "Simple Rating System",
       y = "Count") +
  theme_minimal() +
  geom_vline(aes(xintercept = mean(srs)), col="red")

```

```{r scatters, echo=FALSE, fig.width=12, fig.height=12, fig.cap= "Relationship between Simple Rating System and several adcanced team statistics, coloured by whether a schools winning % was above .500 or not"}
scatter1 <- ncaab_stats %>%
  ggplot(aes(x = pace, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Pace",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter2 <- ncaab_stats %>%
  ggplot(aes(x = o_rtg, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Offensive Rating",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter3 <- ncaab_stats %>%
  ggplot(aes(x = `3pa_rate`, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "3-Point Attempt Rate",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter4 <- ncaab_stats %>%
  ggplot(aes(x = ts_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "True Shooting %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter5 <- ncaab_stats %>%
  ggplot(aes(x = trb_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Total Rebound %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter6 <- ncaab_stats %>%
  ggplot(aes(x = ast_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Assist %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter7 <- ncaab_stats %>%
  ggplot(aes(x = stl_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Steal %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter8 <- ncaab_stats %>%
  ggplot(aes(x = blk_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Block %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter9 <- ncaab_stats %>%
  ggplot(aes(x = tov_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Turnover %",
       y = "Simple Rating System",
       colour = "Over .500"
       )

scatter10 <- ncaab_stats %>%
  ggplot(aes(x = orb_percent, y = srs, colour = over_500)) +
  geom_point()+
  theme_bw() +
  labs(x = "Offensive Rebound %",
       y = "Simple Rating System",
       colour = "Over .500"
       )
ggarrange(scatter1, scatter2, scatter3, scatter4, scatter5, scatter6, scatter7, scatter8, scatter9, scatter10, ncol=3, nrow=4, common.legend = TRUE, legend="bottom")
```

```{r summarytab, fig.cap = "Summary statistics of important team variables"}
ncaab_stats %>% select(srs, o_rtg, ts_percent, trb_percent, tov_percent) %>% datasummary_skim()
```

```{r}
ncaab_stats %>% select(srs, o_rtg, ts_percent, trb_percent, tov_percent) %>% datasummary_correlation()
```
